\documentclass[PhD]{dukethesis2006}
%\usepackage{setspace}
\usepackage{forest}
\usepackage[notes]{biblatex-chicago}
\usepackage{etoolbox}
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
\usepackage{amssymb,amsmath, amsthm,mathtools}
\usepackage{longtable,booktabs}

%\usepackage{newpxmath}
\usepackage{heuristica}
\usepackage[heuristica,vvarbb,bigdelims]{newtxmath}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\renewcommand*\oldstylenums[1]{\textosf{#1}}
\usepackage[left=1.5in,right=1in,top=1.5in,bottom=1.5in]{geometry}
\input{bob_macros.tex}
\bibliography{diss.bib}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\AtBeginEnvironment{quote}{\singlespacing\small}

\newenvironment{dedication}
  {\clearpage           % we want a new page          %% I commented this
   \thispagestyle{empty}% no header and footer
   \vspace*{\stretch{1}}% some space at the top
   \itshape             % the text is in italics
   \raggedleft          % flush to the right margin
  }
  {\par % end the paragraph
   \vspace{\stretch{3}} % space at bottom is three times that at the top
   \clearpage           % finish off the page
  }
\author{Lok C. Chan}
\advisor{Kevin D. Hoover}
\member{Vincent Conitzer}
\member{Felipe De Brigard}
\member{Carlotta Pavese}



\department{Philosophy}
%\subject{xxx} If this is used, "subject" has to be un-commented in the cls file in several places
\title{Deliberative Bayesianism:\\Abduction, Reflection, and the Weight of Evidence}
\begin{document}


\maketitle

\makeabstract
\Copyright
\abstract
In this dissertation, I defend the thesis that an epistemic judgment of
probability must be interpreted against the background of the
\emph{context of inquiry} in which it is made: in the \emph{abductive} context,
judgments of probability are matters of \emph{decision}, made
strategically in service to the investigative goal of the inquirer; in
\emph{deduction}, probabilities are \emph{derived} based on the
premises chosen in abduction, in order to explicate the implied
commitments the agent may incur from those decisions; during the
\emph{inductive} stage, the inquirer is expected to conduct her empirical
investigation in a deliberate manner, in accordance with the assertions and
decisions she made during abduction and deduction, collectively referred
to as the \emph{deliberative context}.

I set the stage by proposing a pragmatist reading of Bas van Fraassen's Reflection Principle and his \emph{voluntarist} interpretation of assertions of degrees of beliefs as performative locutions to express the intention to undertake a proportional epistemic commitment. I argue for a refinement of this view that I call \emph{deliberativism}, which introduces an abductive dimension to understand the normative force that regulates these epistemic judgments. I then argue that decisions made in the context of abduction have inferential repercussions on the validity of inductive inference. In particular, I situate deliberativism in the context of statistical inference by critically examining a problem in the literature called \emph{optional stopping}, which occurs when the experimenter's intention to stop can manipulate the statistical significance of the data. The last chapter explores the Pericean idea of deductive reasoning as the strategic interrogation of a provisionally chosen hypothesis by focusing on J. M. Keynes' notion of \emph{the weight of evidence}.


\begin{dedication}
	To Mom, who fought for my fighting chance
	\end{dedication}

\singlespacing
\tableofcontents
\listoffigures
\listoftables
\doublespacing

\acknowledgements

I would like to thank my mother, but I cannot, as she sadly passed away two years after I started graduate school. Trying to live well, I hope, is the next best thing.

Thankfully, there are people who I would like to and can thank.

Thank you, Elisa, my partner. I think the dissertation was more work for you than me. You made this possible. 

Many thanks to my advisor and supervisor Kevin Hoover---not only for t he immense support and encourage, but also for the countless hours in the Perk chatting about philosophy and working on my ideas. 

Thank you to my committee members, Vince Conitzer, Felipe De Brigard, and Carlotta Pavese. Thank you all for dealing with my fickleness: the dissertation I have now is pretty different than the project I described. I cannot ask for a better committee that can reflect the kind of researcher I aim to be. 

Special thanks to my friends Lauren Bunch, Jonathan Chen, and Randy Mayes, for reading my work and catching my increasingly innovative ways of making typos, which, like a virus that develops resistance to antibiotic, have developed a strong immunity to proofreading.

Lastly, without the awesome people who helped me in the philosophy department, I probably would not have survived this last few years. I want to especially thank Gopal Sreenivasan and Janelle Haynes for guiding me through this process. I also need to thank Owen Flanagan and David Wong, for the kindness that both of you have shown during a very difficult time in my life. 

\include{intro}

\include{reflectionpragmatist}

\include{stoppingrule}
\include{weight}
\include{conclusion}

\printbibliography[heading=bibintoc]

\biography

	Lok Chan was born in Hong Kong in 1984. He studied classical guitar performance in California State University, Sacramento. When his loving mother advised him to study something more practical and lucrative in nature, he decided to pursue as a second major philosophy with an emphasis on logic and philosophy of science. He then went on to complete a Master's degree in philosophy from San Francisco State University. At Duke University, he received the Bass Instructional Fellowship, the Graduate Student Training Enhancement Grant, Professional Development Grant and the Summer Research Fellowship. He was a Lab Fellow at Franklin Humanities Institute's Lab of Digital Knowledge, and a member of the Moral AI group, a Bass Connections team. He served on the Versatile Humanists at Duke advisory board. He will work as a postdoctoral researcher at Duke's Department of Population Health Sciences and Social Sciences Research Institute.

\end{document}